---
  title: 'Figures'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
  code_folding: show
toc: true
toc_depth: 3
toc_float: yes
number_sections: false
theme: cerulean
highlight: haddock
includes: 
  in_header: '../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---
  
```{r}

library(MyFunctions)

my_lib(c(
  "tidyverse",
  "sf",
  "here",
  "raster",
  "sp",
  "cowplot"
))


# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select

# library(tidyr) Erase
# library(dplyr) Erase
# library(raster)
library(RColorBrewer)
library(rgdal)
# library(sf)
library(gstat)
library(stringr)
library(fields)
library(cowplot)
library(beyonce)
# library(here)

# load spatial files (ocean raster and regions shapefile)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/gh-pages/src/R/spatial_common.R")

regions_shape <- as(regions, "Spatial")
land <- regions_shape[regions_shape$rgn_type %in% c("land", "land-disputed", "land-noeez", "eez-inland"), ]

chi_breaks <-  c(-1, 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.25, 1.5, 1.75, 2, 2.5, 3, 3.5, 4.0, 5, 100)
chi_cols <- c("#9E0142", "#B91F48", "#D53E4F", "#F46D43", "#FDAE61", "#FEE08B", "#FFFFBF", "#EFF9FF", "#BDE4FC", "#3288BD")
chi_cols = rev(colorRampPalette(chi_cols)(length(chi_breaks)-1)) 
chi_legend_labels <- c(0, 0.2, 0.6, 1, 1.5, 2, 3, 4.0, ">5")
chi_label_sequence <- c(1, 3,  5,   7, 9,  11, 13, 15, 17)

```

# Map of CHI

This is figure 1 B

```{r chi_map}

## function to make plots

legend.shrink <- 0.7
legend.width <- 0.7


source(here("ms_figs/ncba_chi_plot_function.R"))

for(year in 2003:2013){ #year = 2013
  # png(sprintf("paper/figures/chi_maps/chi_%s.png", year), res=500, width=6, height=3, units="in")  
    png(sprintf("paper/figures/chi_maps/chi_%s.png", year), res=500, width=6, height=3, units="in")
  
  # p_rast <- raster(file.path(dir_M, sprintf("git-annex/impact_acceleration/impact/cumulative_impact/chi_%s.tif", year)))
  p_rast <- raster(file.path(dir_M, sprintf("~/Library/CloudStorage/OneDrive-UBC/Data/ncba_stressor_nam/Data/chi/data//chi_%s.tif", year)))
                
  chi_plot(raster_data=p_rast,  title="", 
                cols=chi_cols, color_breaks = chi_breaks,
                legend_break_labels = chi_legend_labels,
           label_sequence = chi_label_sequence)
  dev.off()
}


```

# Plot map of trend

```{r trend_map}

slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
overlay_rast <- raster(file.path(dir_M,  'git-annex/impact_acceleration/impact/trend/sig_overlay.tif'))

#source(here("paper/figures/chi_plot_function.R"))

png("paper/figures/trend/chi_trend.png", res=500, width=6, height=3, units="in")  
trend_plot(plotRaster=slope, overlay=TRUE, overlay_rast=overlay_rast)
dev.off()

png("paper/figures/trend/chi_trend_no_overlay.png", res=500, width=6, height=3, units="in")  
trend_plot(plotRaster=slope, overlay=FALSE)
dev.off()


```



## EEZ CHI data exploration

data used `eez_chi.csv`

```{r}

eez_chi <- my_path("D","chi","eez_chi.csv", read = T)

# north_america <- 
  eez_chi %>% 
  filter(rgn_id %in% c(116,135,163,218)) %>% 
    group_by(rgn_name) %>% 
    summarise(
      mean_eez_chi = mean(value, na.rm = T)
    ) %>% 
    ggplot() +
    geom_point(
      aes(
        x = reorder(rgn_name,mean_eez_chi),
        y = mean_eez_chi,
        color = rgn_name
      )
    )

```


## CHI trend Figure 1A and 1B

### Get data
data used `cumulative_impact_trend_2003-2013.tif` from knb

```{r}

# Get slope data from Halpern et al (2019)
slope <- raster(my_path("D","chi","cumulative_impact_trend_2003-2013.tif"))
# Plot to make sure it is the correct data
# trend_plot(plotRaster=slope, overlay=FALSE)


# CHI for 2013
chi_2013 <- raster(my_path("D","chi","cumulative_impact_2013.tif"))

# Load SAU EEZs we are interested in
# MyFunctions::my_sf("SAU") %>% arrange(name) %>% pull(name) %>% unique() # Check eez names
sau_sf <- MyFunctions::my_sf("SAU") %>%
  filter(name %in% c(
    "USA (East Coast)","USA (West Coast)","USA (Gulf of Mexico)","USA (Alaska, Arctic)","USA (Alaska, Subarctic)",
    "Hawaii Main Islands (USA)","Hawaii Northwest Islands (USA)","Puerto Rico (USA)" ,"US Virgin Islands","USA (Alaska, Arctic)",
    "Mexico (Pacific)","Mexico (Atlantic)",
    "Canada (Arctic)","Canada (East Coast)","Canada (Pacific)")) %>%
  rename(sf_eez_name = name) %>% 
  left_join(my_path("Spa","SAU","sau_matching_names.csv", read = T)) %>% 
  select(sf_eez_name, name = un_country,eezid) %>% 
  st_transform(crs(slope))

# Get land countries
# rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>% arrange(admin) %>% pull(admin) %>% unique()
world_land <- rnaturalearth::ne_countries(scale = 'large', returnclass = c("sf")) %>% 
  st_transform(4326) %>% 
  filter(admin %in% c("Canada","United States of America","Puerto Rico","Mexico")) %>% 
  st_shift_longitude()

# Get SAU codes for matching data
sau_codes <- my_data("sau_index") %>% 
  dplyr::select(x = 1, everything()) %>% 
  filter(x %in% sau_sf$eezid)

# Load SAU as grid to extract data
# Transform SAU grid to a points sf
sau_sf <- st_as_sf(my_data("dbem_coords"),
                    coords = c("lon", "lat"), 
                    crs = 4326) %>% 
  st_transform(crs(slope)) %>% 
  filter(index %in% sau_codes$index)
  

# Ohara's color pallet for figure 2C
### diverging color palette from ColorBrewer
div_pal <- c('#8e0152','#c51b7d','#de77ae','#f1b6da','#fde0ef',
             '#f7f7f7', ### mid color
             '#e6f5d0','#b8e186','#7fbc41','#4d9221','#276419') %>% rev()



```


### Incorporate data into SAU shapefile

```{r}

# Extract slope raster values at grid points of the SAU SF
sau_sf$chi_value <- extract(slope, sau_sf)

# Extract chi 2013 raster values at grid points of the SAU SF
sau_sf$chi_2013 <- extract(chi_2013, sau_sf)

# Final details of shapefile projection
sau_chi_sf <- sau_sf %>% 
  # Filter the region we want
  st_transform(4326) %>%
  st_shift_longitude()


```

### Map it

#### Slope Figure

```{r}
slope_map <-
  ggplot() +
  geom_sf(data = sau_chi_sf, aes(color = chi_value), size = 0.3) +
  geom_sf(data = world_land, aes(), color = "black") +
  scale_color_gradient2(
    # "Anual change in\ncumulative impact",
    low = "#3B9AB2", mid = "white", high = "#F21A00",
    breaks = c(-0.2,-0.1,0,0.1,0.2),
    labels = c(-0.2,-0.1,0,0.1,0.2),
    na.value = 'grey80') +
  theme_classic() +
  my_ggtheme_p(ax_tx_s = 8,ax_tl_s = 10,leg_tl_s = 10,leg_tx_s = 8,facet_tx_s = 10,leg_pos = "right")  +
    theme(plot.margin = unit(c(.05, 0, .05, 0), units = 'cm'),
        # legend.background = element_blank(),
        legend.key.width = unit(.25, 'cm'),
        # axis.text.x = element_blank(),
        legend.title = element_blank()
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

slope_map

```


#### CHI Figure

```{r}

pal <- wesanderson::wes_palette("Zissou1", 100, type = "continuous")

chi_map <-
  ggplot() +
  geom_sf(data = sau_chi_sf, aes(color = chi_2013), size = 0.3) +
  # geom_sf(data = sau_chi_sf %>% mutate(chi_2013 = ifelse(chi_2013 >4,4,chi_2013)), aes(color = chi_2013), size = 0.3) +
  geom_sf(data = world_land, aes(), color = "black") +
    # scale_color_gradientn(
    #   "Cumulative\nhuman impact",
    #   colours = pal,
    #   breaks = seq(0,5,1),
    #   labels = seq(0,5,1),
    #   limits = c(0,5)
    #   ) + 
    scale_color_viridis_c(
      # "Cumulative\nhuman impact",
      breaks = seq(0,5,1),
      labels = seq(0,5,1),
      limits = c(0,5)
      # breaks = seq(0,4,1),
      # labels = c(seq(0,3,1),">4"),
      # limits = c(0,4)
      ) + 
  theme_classic() +
  my_ggtheme_p(ax_tx_s = 8,ax_tl_s = 10,leg_tl_s = 10,leg_tx_s = 8,facet_tx_s = 10,leg_pos = "right")  +
    theme(plot.margin = unit(c(.05, 0, .05, 0), units = 'cm'),
        # legend.background = element_blank(),
        legend.key.width = unit(.25, 'cm'),
        axis.text.x = element_blank(),
        legend.title = element_blank()
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))

chi_map

```


## Panel 

```{r}
chi_panel <-
  ggdraw() +
    draw_plot(chi_map,  x = .0001, y = 0.5, width = 0.90, height = .49) +
    draw_plot(slope_map,  x = .01, y = 0, width = 0.90, height = .49) +
    draw_label('A', x = .04, y = .98, vjust = 0, size = 10, color = 'black') +
    draw_label('B', x = .04, y = .50, vjust = 0, size = 10, color = 'black') +
    draw_label('Cumulative human impact\n(2013)', x = .94, y = .75, hjust = 0.5, vjust = 1,
               size = 8, color = 'black', angle = 270) +
    draw_label("Anual change in cumulative impact\n(2003-2013)", x = .95, y = .25, hjust = 0.5, vjust = 1,
               size = 8, color = 'black',angle = 270) +
    # draw_label('Intensification\n(percent)', x = .91, y = .18, hjust = 0.5, vjust = 1, 
    #            size = 8, color = 'black', angle = 270) + 
  theme(plot.background = element_rect(fill="white", color = NA))
    
  ggsave(plot = chi_panel, filename = "ncba_chi_map.png",
         height = 20, width = 15, units = 'cm', dpi = 300)
  
  print(chi_panel)

```




#Original Code below











# region zooms

Project to lat long.

```{r zoom_data}

## project raster to lat/long

# make a lat/long raster to copy
sim_rast <- raster(ncol=40000, nrow=20000)

chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))
projectRaster(chi, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/paper/figures/chi_lat_long_2013.tif"), progress="text", overwrite=TRUE)
chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/paper/figures/chi_lat_long_2013.tif"))
plot(chi)

extremes <- raster(file.path(dir_M, 'git-annex/impact_acceleration/paper/figures/chi_slope_extremes.tif'))
projectRaster(extremes, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/paper/figures/extremes_lat_long.tif"), progress="text", overwrite=TRUE)

slope <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/chi_slope.tif'))
projectRaster(slope, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/paper/figures/slope_lat_long.tif"), progress="text", overwrite=TRUE)

slope_overlay <- raster(file.path(dir_M, 'git-annex/impact_acceleration/impact/trend/sig_overlay.tif'))
projectRaster(slope_overlay, sim_rast, method="ngb", filename=file.path(dir_M, "git-annex/impact_acceleration/paper/figures/sig_overlay_lat_long.tif"), progress="text", overwrite=TRUE)

## impacts
## (starting with 2013 files)
library(doParallel)
library(foreach)
library(parallel)


impact_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/stressor_impact"), recursive = TRUE, 
                           pattern = ".tif", full="TRUE")

impacts_2013 <- grep("2013", impact_files, value=TRUE) 
impacts_2013 <- grep("uv|benthic_str", impacts_2013, value=TRUE, invert=TRUE)

registerDoParallel(8)

foreach(impact = impacts_2013,.packages="dplyr") %dopar%{
  #for(impact in impacts_2013){ # impact = impacts_2013[2]
  impact_name <- paste("latlong", basename(impact), sep="_")
  
  imp_rast_mol <- raster(impact)
  
  raster::projectRaster(imp_rast_mol, sim_rast, method="ngb", over=TRUE, filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/paper/figures/impacts_lat_long/%s", impact_name)), overwrite=TRUE)
}

check <- raster("/home/shares/ohi/git-annex/impact_acceleration/paper/figures/impacts_lat_long/latlong_dem_dest_2013.tif")

### Trend data

trend_files <- list.files(file.path(dir_M, "git-annex/impact_acceleration/impact/trend/impacts"), 
                          pattern = ".tif", full="TRUE")

registerDoParallel(8)

foreach(trend = trend_files,.packages="dplyr") %dopar%{
  #for(impact in impacts_2013){ # trend = trend_files[2]
  trend_name <- paste("latlong", basename(trend), sep="_")
  
  imp_rast_mol <- raster(trend)
  
  raster::projectRaster(imp_rast_mol, sim_rast, method="ngb", over=TRUE, filename=file.path(dir_M, sprintf("git-annex/impact_acceleration/paper/figures/trend_lat_long/%s", trend_name)), overwrite=TRUE)
}

check <- raster("/home/shares/ohi/git-annex/impact_acceleration/paper/figures/trend_lat_long/latlong_dem_dest_2013.tif")


```

Figure out regions.
```{r zoom_regions}

#regions_wgs <- st_transform(regions, 4326)
land_wgs <- spTransform(land, CRS("+init=epsg:4326"))

chi_slope <- raster(file.path(dir_M, "git-annex/impact_acceleration/paper/figures/extremes_lat_long.tif"))

plot(chi_slope)
tmp <- zoom(chi_slope)

# low fast (not really any of these!)
#mad <- extent(45, 49, -17, -13)

###
# high decrease
ns <- extent(-11, 0, 55, 60)

# high increase
#gomaine <- extent(-72, -65, 41, 46)
aus <- extent(112, 119, -36, -31)

# low/decrease
#nz <- extent(164, 174.5, -48, -38)
ak <- extent(-169, -161, 59, 64) 

```

### region: chi plots

```{r zoom chi}

chi_crop <- function(raster_data, crop_rgn, myBreaks, cols, filename){
  png(sprintf("paper/figures/rgn_maps/%s.png", filename), res=600, width=6, height=5, units="in")  
  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))
  
  master <- raster_data
  master_crop <- crop(master, crop_rgn)
  
  plot(c(crop_rgn[1], crop_rgn[2]), c(crop_rgn[3], crop_rgn[4]), 
       type="n", xaxs="i", yaxs="i", axes=FALSE)
  plot(master_crop, col=cols,  breaks=myBreaks, legend=FALSE, add=TRUE)
  plot(land_wgs, add=TRUE, border="gray80", col="gray90", lwd=0.5)
  box("plot", col="gray")
  
  dev.off()
}


chi <- raster(file.path(dir_M, "git-annex/impact_acceleration/paper/figures/chi_lat_long_2013.tif"))


chi_crop(raster_data=chi,  crop_rgn=ak,
                   myBreaks=chi_breaks, cols=chi_cols, filename="chi_ak")
chi_crop(raster_data=chi,  crop_rgn=aus,
                   myBreaks=chi_breaks, cols=chi_cols, filename="chi_aus")
chi_crop(raster_data=chi,  crop_rgn=ns,
                   myBreaks=chi_breaks, cols=chi_cols, filename="chi_ns")

```

### region trend plots

```{r zoom trend}


chi_slope_plot <- function(raster_data, crop_rgn, saveFile, cols){
  
  ## crop region
  master_crop <- crop(raster_data, crop_rgn)
  
  png(sprintf("paper/figures/rgn_maps/%s.png",  saveFile), 
      res=500, width=6, height=5, units="in")  
  par(mar=c(2,2,2,2)) # bottom, left, top, and right
  par(oma=c(0,0,0,0))
  
  plot(c(crop_rgn[1], crop_rgn[2]), c(crop_rgn[3], crop_rgn[4]), 
       type="n", xaxs="i", yaxs="i", axes=FALSE)
  
  low <- rev(c("#E0F3F8", "#ABD9E9", "#74ADD1", "#5C91C2", "#4575B4"))
  high <- rev(c("#A50026", "#C11B26", "#D02926", "#E14631", "#F46D43", "#F9A669", "#FEE090", "#FFF1CC"))

  quants <- quantile(raster_data, c(0.00001, 0.99999))
  low_breaks <- c(minValue(raster_data)-0.01, 
              seq(quants[[1]], 0, by=0.005))
  high_breaks <- c(seq(0, quants[[2]], by=0.005),
                   maxValue(raster_data)+0.01)
  
  low_cols <- colorRampPalette(low)(length(low_breaks)-1)
  high_cols <- colorRampPalette(high)(length(high_breaks)-1)
  cols <- c(low_cols, "#F4FBFC", "#F4FBFC", high_cols)
  

  plot(master_crop, col=cols,  
      breaks=c(low_breaks, high_breaks),
       legend=FALSE, add=TRUE)
  
  plot(land_wgs, add=TRUE, border="gray75", col="gray90", lwd=0.5)
  box("plot", col="gray")
  dev.off()
}

slope <- raster(file.path(dir_M, "git-annex/impact_acceleration/paper/figures/slope_lat_long.tif"))


chi_slope_plot(raster_data=slope, crop_rgn=aus, saveFile="slope_aus", cols=cols)
chi_slope_plot(raster_data=slope, crop_rgn=ns, saveFile="slope_ns", cols=cols)
chi_slope_plot(raster_data=slope, crop_rgn=ak, saveFile="slope_ak", cols=cols)

```


### combine images

```{r zoom_combine}
library(cowplot)
library(magick)

chi1 <- ggdraw() + draw_image("paper/figures/rgn_maps/chi_ak.png")
chi2 <- ggdraw() + draw_image("paper/figures/rgn_maps/chi_ns.png")
chi3 <- ggdraw() + draw_image("paper/figures/rgn_maps/chi_aus.png")


chi_plots <- plot_grid(chi1, chi2, chi3, nrow=1)
save_plot("paper/figures/rgn_maps/chi.png", chi_plots, ncol=1, base_height=1, base_width=4)


trend1 <- ggdraw() + draw_image("paper/figures/rgn_maps/slope_ak.png")
trend2 <- ggdraw() + draw_image("paper/figures/rgn_maps/slope_ns.png")
trend3 <- ggdraw() + draw_image("paper/figures/rgn_maps/slope_aus.png")



trend_plots <- plot_grid(trend1, trend2, trend3, nrow=1)
save_plot("paper/figures/rgn_maps/trend.png", trend_plots, ncol=1, base_height=1, base_width=4)

```



#### Make map showing regions

```{r zoom_global_loc}

# function to convert lat/long extents to moll points and add to plot
plotwgs <- function(x){ # x = aus
  long <- mean(c(x[1], x[2]))
  lat <- mean(c(x[3], x[4]))
  x_sp <- data.frame(lat=lat, long=long)
  coordinates(x_sp) <- c("long", "lat")
  proj4string(x_sp) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
  x_mol <- spTransform(x_sp, CRS(proj4string(land)))
  plot(x_mol, col="red", add=TRUE, pch=16, cex=.7)
}


legend.shrink <- 0.7
legend.width <- 0.7

cols_trans = paste0(chi_cols, "80")

p_rast <- raster(file.path(dir_M,  "git-annex/impact_acceleration/impact/cumulative_impact/chi_2013.tif"))

png("paper/figures/rgn_maps/rgn_locations.png", 
    res=500, width=6, height=3, units="in")  
par(mar=c(1,1,1,1)) # bottom, left, top, and right
par(oma=c(0,0,0,0))
plot(p_rast, col=cols_trans, axes=FALSE, box=FALSE, breaks=chi_breaks, legend=FALSE)

# add axis with fields package function:
  # add axis with fields package function:
  break_locations <- seq(0, length(chi_breaks), length.out=length(chi_breaks)) # breaks for colors for legend
  legend_label_locations <- break_locations[seq(1, length(break_locations), 2)] # label locations (every other color labeled)
  
  fields::image.plot(p_rast, #zlim = c(min(myBreaks), max(myBreaks)), 
                     legend.only = TRUE, 
                     legend.shrink=legend.shrink,
                     legend.width=legend.width,
                     col = cols_trans,
                     #legend.lab=title_legend,
                     breaks=break_locations,
                     axis.args=list(cex.axis=0.6, at=legend_label_locations, labels=chi_legend_labels))
plot(land, add=TRUE, border="gray80", col="gray90", lwd=0.5)

plotwgs(ns)
plotwgs(aus)
plotwgs(ak)

dev.off()

```


# Trend for each 3nm eez impact circleplot

```{r 3nm_eez_trend_circleplot}

# UN georegions
un_rgn <- UNgeorgn_nm %>%
  dplyr::select(rgn_id, georegion = r1_label)

# CHI for each region
chi <- read.csv("paper/zonal_data_eez/eez_chi_3nm_trend.csv") %>%
  dplyr::select(rgn_id, rgn_name, chi=value) %>%
  left_join(un_rgn, by="rgn_id")


# get georegion ranks based on mean chi values, and use to rank chi data
rank_georgn <- chi %>%
  group_by(georegion) %>%
  summarize(mean_chi = mean(chi),
            count = length(chi)) %>%
  arrange(mean_chi)

chi$georegion <- factor(chi$georegion, levels=rank_georgn$georegion)


### Organize the chi data and add a few extra variables to help plotting
# Including empty spaces to add a y-axes

rank_rgn <- chi %>%
  arrange(georegion, chi)

# add empty space
empty_bar <- 6
to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn)) )
colnames(to_add) = colnames(rank_rgn)
to_add$rgn_name <- as.character(1:empty_bar)
rank_rgn  <- rbind(to_add, rank_rgn)

# modify region names to be shorter
rank_rgn <- rank_rgn %>%
  mutate(rgn_name_short = rgn_name, 
         rgn_name_short = gsub("Islands", "Isl", rgn_name_short),
         rgn_name_short = gsub("Island", "Isl", rgn_name_short),
         rgn_name_short = gsub("Democratic", "Dem", rgn_name_short),
         rgn_name_short = gsub("Republic", "Rep", rgn_name_short),
         rgn_name_short = gsub("South", "S", rgn_name_short),
         rgn_name_short = gsub("American", "Am", rgn_name_short),
         rgn_name_short = gsub("the United States", "US", rgn_name_short),
         rgn_name_short = gsub("Territory", "Terr", rgn_name_short),
         rgn_name_short = gsub("Saint", "St", rgn_name_short),
         rgn_name_short = gsub(" and ", " & ", rgn_name_short),
         rgn_name_short = gsub("Republique", "Rep", rgn_name_short),
         rgn_name_short = gsub("Dem Rep of the", "Dem Rep of", rgn_name_short),
         rgn_name_short = gsub("Georgia and the", "Georgia and", rgn_name_short),
         rgn_name_short = gsub("St Vincent and the", "St Vincent and", rgn_name_short),
         rgn_name_short = gsub("Howland Isl & Baker Isl", "Howland & Baker Isl", rgn_name_short),
         rgn_name_short = gsub("Northern", "N", rgn_name_short), 
         rgn_name_short = gsub("Reunion", "Reunion   ", rgn_name_short))


# some code to orient the country labels
sequence_length = length(unique(rank_rgn$rgn_name))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)

rank_rgn$angle <- c(first_angles,second_angles)
rank_rgn$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))

# color for region labels
rank_rgn <- rank_rgn %>%
  mutate(rgn_name = factor(rgn_name, unique(rgn_name))) %>%
  mutate(georegion = factor(georegion, unique(georegion))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(chi), "white", color))


########### marking georegions

# add column to identify when georegion changes
rgn_shift <- rank_rgn %>%
  mutate(georegion = ifelse(is.na(georegion), "tmp", georegion)) %>%
  mutate(georegion = as.factor(georegion)) %>%
  mutate(region_shift = as.numeric(georegion) - lag(as.numeric(georegion)), default=first(as.numeric(georegion)))

rgn_shift <- which(rgn_shift$region_shift > 0)
rgn_shift <- c(1, rgn_shift) -0.5
rgn_shift <- data.frame(rgn_shift_x=rgn_shift,
                        georegion = rank_georgn$georegion,
                        name_x= c(25, 51, 68, 90, 110, 145, 200), # 140
                        name_y=c(0.1, 0.1, 0.1, 0.12, 0.11, 0.1, 0.11))
rgn_shift <- rgn_shift %>%
  mutate(georegion = as.character(georegion)) %>%
  mutate(georegion = ifelse(georegion == "Latin America and the Caribbean", "Latin America and\nthe Caribbean",
                            georegion))

# get impact data
impacts <- read.csv("paper/zonal_data_eez/eez_3nm_impacts_trend.csv") %>%
  left_join(un_rgn)

# ## add some blanks to separate groups
to_add <-  data.frame( matrix(NA, empty_bar*nlevels(as.factor(impacts$pressure)), ncol(impacts)) )
colnames(to_add) <- colnames(impacts)
to_add$pressure <- rep(levels(as.factor(impacts$pressure)), each=empty_bar)
to_add$value <-  0
to_add$rgn_name <- as.character(rep(1:empty_bar, nlevels(as.factor(impacts$pressure)))) 

impacts <- rbind(to_add, impacts, to_add)


## rank of pressure categories
impacts %>%
  group_by(pressure) %>%
  summarize(mean = mean(value)) %>%
  arrange(mean)

pressure_name <- data.frame(pressure = c("sst", "slr", "oa", 
                                         "shipping",
                                         "light", "nutrient", "direct_human", "organic",
                                         "pel_hb", "dem_nondest_hb","pel_lb", 
                                         "dem_nondest_lb", "dem_dest", "art_fish"),
                            pressure_name = c("sst", "slr", "oa",
                                              "shipping", "light pollution", 
                                              "nutrient pollution", "direct human", "organic pollution",  
                                              "comm fish: pel hb", "comm fish: dem nondest hb", "comm fish: pel lb", 
                                              "comm fish: dem nondest lb", "comm fish: dem dest", "artisanal fishing"))

impacts <- impacts %>%
  left_join(pressure_name, by = "pressure")

impacts$pressure_name <- factor(impacts$pressure_name, levels=rev(pressure_name$pressure_name))
impacts$rgn_name <- factor(impacts$rgn_name, levels=unique(rank_rgn$rgn_name))


library(beyonce)

myPalette <- c(beyonce_palette(18, 15, type=c("continuous"))[1:6],
               beyonce_palette(18, 25, type=c("continuous"))[15:18],
               beyonce_palette(18, 15, type=c("continuous"))[8],
               beyonce_palette(18, 20, type=c("continuous"))[16:19])


## some theme stuff to make the circle plot look nice

circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                      legend.position="none",
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank())

# circle plot
# https://www.r-graph-gallery.com/299-circular-stacked-barplot/
p <- ggplot(data=impacts, aes(x=rgn_name, y=value, fill=pressure_name)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.13, ymax=0.25), alpha=0)  +
  geom_text(data=rank_rgn, aes(x=rgn_name, y=0.16, label=rgn_name_short, angle=angle, hjust=hjust, color=color), inherit.aes = FALSE) +
  geom_segment(x = 5, y = 0, xend = dim(rank_rgn)[1]+1, yend = 0, colour = "black", alpha=1, size=0.5) +
  geom_segment(x = 7, y = -0.05, xend = dim(rank_rgn)[1]-1, yend = -0.05, colour = "gray", alpha=1, size=0.5) +
  annotate("text", x = c(3,3,3,3), y = c(-0.05, 0, 0.05, 0.1), label = c(-0.05, 0, 0.05, 0.1), color="darkgrey", angle=-8, size=4) +
  scale_colour_identity() +
  scale_fill_manual(values=myPalette) +
  coord_polar() +
  geom_segment(data = rgn_shift, aes(x=rgn_shift_x, xend=rgn_shift_x, y=rep(-0.05, dim(rgn_shift)[1]), yend=rep(0.2, dim(rgn_shift)[1])), colour="gray", size=0.5, inherit.aes=FALSE) +
  geom_text(data=rgn_shift, aes(x=name_x, y=name_y, label=georegion), inherit.aes=FALSE, size=5) +
  #geom_point(data=chi, aes(y=chi, x=rgn_name), fill="black", shape="|", size=2.5) +
  circle_theme

p
ggsave('paper/figures/trend/eez_3nm_trend_circleplot.jpg', height=18, width=18, units=c("in"))

```


### Habitat/impact heat map

```{r hab_impact_heat_map}

## get habitat names that are readable
habitat_nms <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name = plot_name3)

### get impact data for each habitat
hab_imp <- read.csv("paper/zonal_data_habitat/habitat_pressure_2013.csv") %>%
  left_join(habitat_nms, by="habitat") 

### rank of habitats for ordering
total_hab_data <- hab_imp %>%
  group_by(name) %>%
  summarize(impact_on_hab=sum(mean)) %>%
  arrange(impact_on_hab)


## rank of pressure categories for ordering 
## (although the categories weren't super coordinated)
hab_imp %>%
  group_by(impact) %>%
  summarize(mean = mean(mean)) %>%
  arrange(mean)

impact_name <- data.frame(impact = c("sst", "oa", "slr", 
                                     "shipping",
                                     "nutrient", "organic", "direct_human", "light",
                                     "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb",
                                     "dem_nondest_lb", "art_fish"),
                          impact_name = c("sst", "oa", "slr", 
                                          "shipping", 
                                          "nutrient pollution", "organic pollution", "direct human", "light pollution",
                                          "comm fish: pel hb", "comm fish: pel lb", "comm fish: dem dest", 
                                          "comm fish: dem nondest hb", "comm fish: dem nondest lb",
                                          "artisanal fishing"))

hab_imp <- hab_imp %>%
  left_join(impact_name, by = "impact")

# order factors for plotting
hab_imp$name <- factor(hab_imp$name, 
                       levels = total_hab_data$name)

hab_imp$impact_name <- factor(hab_imp$impact_name, levels=rev(impact_name$impact_name))

hab_cols <- rev(c("#9E0142", "#B91F48", "#D53E4F", "#F46D43", "#FDAE61", "#FEE08B", "#FFFFBF", "#EFF9FF", "#3288bd"))

# saving this file for a ben request: Jan 17 2021
write_csv(hab_imp, "habitats/habitat_stressor_impact_Jan17_2021.csv")
### done with request....

# plot the data
tmp <- ggplot(hab_imp, aes(y=impact_name, x=name)) +
  geom_tile(aes(fill=mean, 
                text=sprintf("habitat: %s \nimpact: %s \ntotal impact: %s" , name, impact_name, round(mean, 2))), color="white") +
  scale_fill_gradientn(colors=hab_cols, name="Impact") +
  ylab("") +
  xlab("") +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=11),
        axis.text.y=element_text(size=11),
        plot.margin=unit(c(1,1,1,1.5),"cm")) 
tmp
ggsave("paper/figures/habitat/carpet_plot_habitat_impact.png", width=10, height=5, units=c("in"), dpi=300)

```

## Habitat Trend and CHI scatterplot
```{r}

filenames <- list.files("paper/zonal_data_habitat", pattern="impacts.csv", full=TRUE)
datalist = lapply(filenames, function(x){read.csv(file=x,header=T)})

hab_chi <- bind_rows(datalist) %>%
  filter(year==2013)

hab_trends <- list.files("paper/zonal_data_habitat", pattern="trend.csv", full=TRUE)
datalist = lapply(hab_trends, function(x){read.csv(file=x,header=T)})

hab_data <- bind_rows(datalist) %>%
  group_by(habitat) %>%
  summarize(trend_total=sum(trend)) %>%
  left_join(hab_chi)

hab_names <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name=plot_name3)

# combine habitats with same raster locations
hab_data <- hab_data %>%
  left_join(hab_names) %>%
  rename(trend = trend_total,
         chi = impact) %>%
  filter(habitat != "ice")


library(ggrepel)
library(cowplot)
ggplot(hab_data, aes(x=chi, y=trend)) +
  geom_point(size=5, alpha=0.5) +
  geom_text_repel(aes(label = name),
                  # box.padding   = 0.35, 
                  #  point.padding = 0.5,
                  segment.color = 'grey50',
                  segment.size = 0.2,
                  point.padding = 0.2,
                  size = 4) +
  labs(x = "Cumulative human impact, 2013", y = "Trend, cumulative human impact")

ggsave("paper/figures/avg_chi_vs_trend_habitat.png", width=8, height=5, units=c("in"), dpi=300)
write_csv(hab_data, "paper/tables/chi_trend_2013.csv")
```



# Trend for each habitat impact circleplot

```{r}

hab_trends <- list.files("paper/zonal_data_habitat", pattern="trend.csv", full=TRUE)
datalist = lapply(hab_trends, function(x){read.csv(file=x,header=T)})

hab_trend_df <- bind_rows(datalist)

trend_data <- read.csv("habitats/habitat.csv") %>%
  dplyr::select(habitat=tif_name, name=plot_name3) %>%
  left_join(hab_trend_df) %>%
  filter(name != "ice edge")


# determine which habitats have the largest/smallest trends:
total_trend <- trend_data %>%
  group_by(name) %>%
  summarize(trend_total = sum(trend)) %>%
  arrange(trend_total) 


# some code to create a naming file that 
# orients the country labels
# and adds a gap in the circle plot for a axis label

empty_bar <- 1
to_add <- data.frame(matrix(NA, empty_bar, ncol(total_trend)) )
colnames(to_add) = colnames(total_trend)
to_add$name <- as.character(1:empty_bar)
total_trend  <- rbind(to_add, total_trend)

sequence_length = dim(total_trend)[1]
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 160/length(first_sequence) * first_sequence)
second_angles = c(-90 - 185/length(second_sequence) * second_sequence)

total_trend$angle <- c(first_angles,second_angles)
total_trend$hjust <- c(rep(0, length(first_sequence)), 
                       rep(1, length(second_sequence)))

total_trend <- total_trend %>%
  mutate(name = factor(name, unique(name))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(trend_total), "white", color))

## add empty data rows to full trend data here
to_add <-  data.frame( matrix(NA, empty_bar*nlevels(as.factor(trend_data$impact)), ncol(trend_data)))
colnames(to_add) <- colnames(trend_data)
to_add$trend <-  0
to_add$impact <- rep(unique(trend_data$impact), empty_bar)
to_add$name <- as.character(rep(1:empty_bar, nlevels(as.factor(trend_data$impact)))) 

trend_data <- bind_rows(to_add, data.frame(trend_data))

## rank of pressure categories
trend_data %>%
  group_by(impact) %>%
  summarize(mean = mean(trend)) %>%
  arrange(mean)

impact_name <- data.frame(impact = c("sst", "oa", "slr", 
                                     "shipping",
                                     "nutrient", "organic", "direct_human", "light",
                                     "pel_hb", "pel_lb", "dem_dest", "dem_nondest_hb",
                                     "dem_nondest_lb", "art_fish"),
                          impact_name = c("sst", "oa", "slr", 
                                          "shipping", 
                                          "nutrient pollution", "organic pollution", "direct human", "light pollution",
                                          "comm fish: pel hb", "comm fish: pel lb", "comm fish: dem dest", 
                                          "comm fish: dem nondest hb", "comm fish: dem nondest lb",
                                          "artisanal fishing"))

trend_data <- trend_data %>%
  mutate(impact = gsub("_trend", "", impact)) %>%
  left_join(impact_name, by = "impact")


trend_data$impact_name <- factor(trend_data$impact_name, levels=rev(impact_name$impact_name))
trend_data$name <- factor(trend_data$name, levels=total_trend$name)

library(beyonce)

myPalette <- c(beyonce_palette(18, 15, type=c("continuous"))[1:6],
               beyonce_palette(18, 25, type=c("continuous"))[15:18],
               beyonce_palette(18, 15, type=c("continuous"))[8],
               beyonce_palette(18, 20, type=c("continuous"))[16:19])


## some theme stuff to make the plot look nice
circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                      legend.position="none",
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank())

# circle plot
# https://www.r-graph-gallery.com/299-circular-stacked-barplot/
p <- ggplot(data=trend_data, aes(x=name, y=trend, fill=impact_name)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.04, ymax=0.24), alpha=0)  +
  geom_text(data=total_trend, aes(x=name, y=trend_total+0.01, label=name, angle=angle, hjust=hjust, color=color), size=3, inherit.aes = FALSE) +
  scale_colour_identity() +
  geom_segment(x = 1.5, y = 0, xend = dim(total_trend)[1]+0.5, yend = 0, colour = "black", alpha=1, size=0.3) +
  annotate("text", x = c(1,1,1), y = c(0, 0.05, 0.1), label = c(0, 0.05, 0.1), color="darkgrey", angle=-10, size=3) +
  scale_fill_manual(values=myPalette) + 
  coord_polar() +
  circle_theme 


p

ggsave('paper/figures/trend/habitat_trend_circleplot.jpg', height=6, width=6, units=c("in"))


# getting the legend:
p <- ggplot(data=trend_data, aes(x=name, y=trend, fill=impact_name)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(x = 1, ymin = -0.04, ymax=0.20), alpha=0)  +
  scale_fill_manual(values=myPalette) + 
  coord_polar() +
  geom_hline(yintercept=0, color="black") +
  geom_text(data=total_trend, aes(x=name, y=0.09, label=name, angle=angle+10, hjust=hjust), inherit.aes = FALSE) +
  theme(legend.title=element_blank())

p
leg <- cowplot::get_legend(p)
plot(leg)


ggsave('paper/figures/trend/legend_circleplot.jpg', height=6, width=6, units=c("in"))


```

